An Experimental Evaluation of Hybrid Querying on Vectors (EA\&B)

### 2025-07-23 Revision/response.tex

%%%%%%%%%%%%%%%%%%%%%% D2 %%%%%%%%%%%%%%%%%%%%%%
\myfbox{
	\begin{detailed}
		\textit{Section 4.2.1: It is important to run the experiments on larger datasets, e.g. 100 M vectors, or 1B bigann. Otherwise, differences will not show.
			Also, build time parameters such as Radius and Length which impact the Recall need to be considered.
			Also, approaches for partitioning and merging for building very large indexes need to be considered.
			Please also document whether quantization is included in index build time.
			There is no explanation why Vamana is not included in Figure 7.}
	\end{detailed}
}

\noindent
\textbf{Response:} 
Thank you for your insightful comments!

\textbf{Regarding dataset scale and scalability:}

We conducted experiments on a 100M-scale deep dataset and analyzed the performance of each algorithm on the large dataset. Please refer to MR.R7.

@@

\textbf{Regarding the experimental parameter settings:}

For datasets with recommended parameters provided in the original papers, we directly used those recommended settings. For other datasets, we selected the parameter combinations that performed best through empirical evaluation. All parameters used in the experiments are available in our public code repository.

According to our experimental results, for datasets using the recommended parameters, the algorithm rankings remained unchanged. During the search process, we adjusted the search parameters to obtain different Recall and QPS values. The search parameters typically span a wide range of values to ensure the completeness of the experimental results.

Due to the large number of algorithms and datasets involved, we did not include all parameter settings in the paper because of space limitations. The detailed parameter configurations can be found in our public code repository. (\textbf{Please refer to Section 4.1.4, highlighted in blue in the revised manuscript.})

	\textcolor{blue}{
"In the experiments, for datasets with recommended parameters in the original paper, we use the recommended settings. For datasets without recommended parameters, we conduct experiments with different parameter settings and select the one that performs best for the final experiment.Due to the large number of datasets and algorithms involved, the parameters are not listed individually, but they are provided in the code repository."}

### 2025-07-23 VLDB2026/main.tex
change parameter part

		\caption{Comparison of AF-ANN and RF-ANN search algorithms}
		\small
%		\label{tab:compair_1}
%		\begin{tabular}{|l|l|c|c|c|c|c|c|c|}
%			\hline
%			\textbf{Algorithm} & \textbf{Base} & \textbf{Filter Type} & \textbf{AND} & \textbf{OR} & \textbf{Flexible Attributes} & \textbf{Complex Boolean} & \textbf{Dynamic Insert} & \textbf{Multi Thread} \\
%			\hline
%			NHQ & Graph & C & Y & N& N& N & N & N \\
%			FilteredVamana & Graph & C & N & Y & Y & N & Y & Y \\
%			StitchedVamana & Graph & C & N & Y & Y & N & N & Y  \\
%			CAPS & IVF & C & Y & N & Y & N & Y & Y \\
%			ACORN & Graph & B & Y & Y & Y & Y & Y & Y \\
%			UNG & Graph & B & Y & Y & Y & N & Y & Y \\ 
%			Puck & IVF & B & Y & Y & Y & N & Y & Y \\
%			% ParlayANN & Graph+IVF & B & N & N & Y & N & N & Y \\
%			
%			\hline
%			
%		\end{tabular}
	\label{tab:compair_1}
	\begin{tabular}{|l|l|*{12}{c|}}
		\hline

@@

		\multicolumn{2}{|l|}{\textbf{RF}}             & N & N & N & N & Y & N & N & Y & Y & Y & Y & Y \\ \hline
		\multicolumn{2}{|l|}{\textbf{Dynamic Insert}} & N & Y & N & Y & Y & Y & Y & N & Y & Y & Y & Y \\ \hline
		\multicolumn{2}{|l|}{\textbf{Multi Thread}}   & N & Y & Y & Y & Y & Y & Y & N & Y & Y & Y & Y \\ \hline
		\multicolumn{2}{|l|}{\textbf{Disk Support}}           &N  & Y& N & N & N & N & N & N & N & N & N & N \\ \hline
		\multicolumn{2}{|l|}{\textbf{Base}}           & Graph & Graph & Graph & IVF & Graph & Graph & IVF & Graph & Graph & Graph & Graph & Graph \\ \hline
	\end{tabular}

@@

		\centering
		%	 \setlength{\abovecaptionskip}{0cm}
		%	\setlength{\belowcaptionskip}{0.1cm}
		\footnotesize{
			\begin{minipage}{\linewidth}
				\vspace{0.1cm}
				\textcolor{orange}{\textsuperscript{AF} Attribue-filtering,}
				\textcolor{orange}{\textsuperscript{RF} Range-filtering,}
				\textcolor{orange}{Filtered: FilteredVamana,}
				\textcolor{orange}{Stitched: StitchedVamana,}
				\textcolor{orange}{Win: WinFilter,}
				\textsuperscript{A} Post-filtering, 
				\textsuperscript{B} Pre-filtering, 
				\textsuperscript{C} Simultaneous filtering, 
				\textsuperscript{Y} Support, 
				\textsuperscript{N} Unsupport. 
				\textit{AND} indicates whether the algorithm supports attribute filtering with AND operations. 
				\textit{OR} indicates whether the algorithm supports attribute filtering with OR operations. 
				\textit{Flexible Attributes} indicates whether the number of query attributes can differ from the number of attributes used during index construction. 
				\textit{Complex Boolean} refers to whether the algorithm supports advanced Boolean logic beyond just AND and OR, such as NOT or combinations like (A AND B) OR (C AND NOT D). 
				\textit{Dynamic Insert} indicates whether the algorithm supports the operation of dynamically inserting data points.
				\textit{Multi Thread} indicates whether the algorithm supports multi-threaded search,
				\textcolor{orange}{\textit{Disk Support} indicates whether the algorithm supports disk storage,}
				\textit{Base} indicates that the hybrid query algorithm is an improvement based on a specific ANN search algorithm.
		\end{minipage}}

@@

@@

	\textcolor{cyan}{In contrast, FilteredVamana achieves optimal performance under long-tail and uniform distributions, as these enable construction of balanced label subgraphs with minimal inter-label interference, thus improving search accuracy and efficiency. However, its performance declines under power-law and normal distributions: power-law distributions cause excessive concentration of popular labels and fragmentation of niche labels, disrupting search path connectivity, while normal distributions produce ambiguous graph structures due to label frequency similarities, leading to search divergence. 
	Under the power-law distribution, when the NHQ algorithm constructs an index based on fusion distance, points corresponding to high-frequency attribute values make the graph connections extremely dense, while points corresponding to low-frequency attribute values tend to become isolated due to sparse connections. This extreme imbalance in connectivity severely distorts the graph's structure, ultimately significantly impairing the overall performance of the algorithm.
	The search performance of VBASE exhibits a direct dependence on attribute selectivity. As selectivity decreases, search speed declines. Notably, only uniform distribution maintains sufficient selectivity among the tested cases; the other three distributions consistently show low selectivity levels that significantly degrade the operational efficiency of VBASE.}

@@

		%\vspace{0.1cm}
		\begin{subfigure}[t]{0.507\columnwidth}
			\centering
			\captionsetup{font=small}
			\includegraphics[width=\linewidth]{figures/exp/attribute_multimodel.pdf}
			\caption{\footnotesize MultiModel}
			\label{fig:MultiModel}
		\end{subfigure}	
		\hfill
		\begin{subfigure}[t]{0.48\columnwidth}
			\centering
			\captionsetup{font=small}
			\includegraphics[width=\linewidth]{figures/exp/attribute_multimodel_1.pdf}
			\caption{\footnotesize MultiModel}
			\label{fig:MultiModel}
		\end{subfigure}

		\begin{subfigure}[t]{0.507\columnwidth}
			\centering
			\includegraphics[width=\linewidth]{figures/exp/attribute_85.pdf}
			\caption{\footnotesize Different Computer}
			\label{fig:Different Computer}
		\end{subfigure}
		\hfill
		\begin{subfigure}[t]{0.48\columnwidth}
			\centering
			\includegraphics[width=\linewidth]{figures/exp/attribute_100M.pdf}
			\caption{\footnotesize Big Dataset}
			\label{fig:Big Dataset}
		\end{subfigure}

@@

		%\vspace{0.1cm}
		\begin{subfigure}[t]{0.507\columnwidth}
			\centering
			\captionsetup{font=small}
			\includegraphics[width=\linewidth]{figures/exp/range_multimodel.pdf}
			\caption{\footnotesize MultiModel}
			\label{fig:MultiModel range}
		\end{subfigure}
		\hfill
		\begin{subfigure}[t]{0.48\columnwidth}
			\centering
			\captionsetup{font=small}
			\includegraphics[width=\linewidth]{figures/exp/range_multimodel_1.pdf}
			\caption{\footnotesize MultiModel}
			\label{fig:MultiModel range}
		\end{subfigure}
		\hfill
		\begin{subfigure}[t]{0.507\columnwidth}
			\centering
			\includegraphics[width=\linewidth]{figures/exp/range_85.pdf}
			\caption{\footnotesize Different Computer}
			\label{fig:Different Computer range}
		\end{subfigure}
		\hfill
		\begin{subfigure}[t]{0.48\columnwidth}
			\centering
			\includegraphics[width=\linewidth]{figures/exp/range_deep100M.pdf}
			\caption{\footnotesize Big Dataset}
			\label{fig:Big Dataset range}
		\end{subfigure}

>>>>

		\caption{Comparison of AF-ANN search algorithms}
		\small
%		\label{tab:compair_1}
%		\begin{tabular}{|l|l|c|c|c|c|c|c|c|}
%			\hline
%			\textbf{Algorithm} & \textbf{Base} & \textbf{Filter Type} & \textbf{AND} & \textbf{OR} & \textbf{Flexible Attributes} & \textbf{Complex Boolean} & \textbf{Dynamic Insert} & \textbf{Multi Thread} \\
%			\hline
%			NHQ & Graph & C & Y & N& N& N & N & N \\
%			FilteredVamana & Graph & C & N & Y & Y & N & Y & Y \\
%			StitchedVamana & Graph & C & N & Y & Y & N & N & Y  \\
%			CAPS & IVF & C & Y & N & Y & N & Y & Y \\
%			ACORN & Graph & B & Y & Y & Y & Y & Y & Y \\
%			UNG & Graph & B & Y & Y & Y & N & Y & Y \\ 
%			Puck & IVF & B & Y & Y & Y & N & Y & Y \\
%			% ParlayANN & Graph+IVF & B & N & N & Y & N & N & Y \\
%			
%			\hline
%			
%		\end{tabular}
	\label{tab:compair_1}
	\begin{tabular}{|l|l|*{12}{c|}}
		\hline

@@

		\multicolumn{2}{|l|}{\textbf{RF}}             & N & N & N & N & Y & N & N & Y & Y & Y & Y & Y \\ \hline
		\multicolumn{2}{|l|}{\textbf{Dynamic Insert}} & N & Y & N & Y & Y & Y & Y & N & Y & Y & Y & Y \\ \hline
		\multicolumn{2}{|l|}{\textbf{Multi Thread}}   & N & Y & Y & Y & Y & Y & Y & N & Y & Y & Y & Y \\ \hline
		\multicolumn{2}{|l|}{\textbf{Disk Storage Support}}           &N  & Y& N & N & N & N & N & N & N & N & N & N \\ \hline
		\multicolumn{2}{|l|}{\textbf{Base}}           & Graph & Graph & Graph & IVF & Graph & Graph & IVF & Graph & Graph & Graph & Graph & Graph \\ \hline
	\end{tabular}

@@

		\centering
		%	 \setlength{\abovecaptionskip}{0cm}
		%	\setlength{\belowcaptionskip}{0.1cm}
		\footnotesize{
			\begin{minipage}{\linewidth}
				\vspace{0.1cm}
				\textsuperscript{A} Post-filtering, 
				\textsuperscript{B} Pre-filtering, 
				\textsuperscript{C} Simultaneous filtering, 
				\textsuperscript{Y} Support, 
				\textsuperscript{N} Unsupport. 
				\textit{Base} indicates that the hybrid query algorithm is an improvement based on a specific ANN search algorithm. 
				\textit{AND} indicates whether the algorithm supports attribute filtering with AND operations. 
				\textit{OR} indicates whether the algorithm supports attribute filtering with OR operations. 
				\textit{Flexible Attributes} indicates whether the number of query attributes can differ from the number of attributes used during index construction. 
				\textit{Complex Boolean} refers to whether the algorithm supports advanced Boolean logic beyond just AND and OR, such as NOT or combinations like (A AND B) OR (C AND NOT D). 
				\textit{Dynamic Insert} indicates whether the algorithm supports the operation of dynamically inserting data points.
				\textit{Multi Thread} indicates whether the algorithm supports multi-threaded search.
		\end{minipage}}

@@

	\textcolor{blue}{
	\subsubsection{Parameters}
	In the experiments, for datasets with recommended parameters in the original paper, we use the recommended settings. For datasets without recommended parameters, we conduct experiments with different parameter settings and select the one that performs best for the final experiment.Due to the large number of datasets and algorithms involved, the parameters are not listed individually, but they are provided in the code repository.
}

@@

	In contrast, VBASE exhibits the weakest robustness. As a post-filtering system, it performs significantly better under uniform distributions but degrades under long-tailed, normal, and power-law distributions. This degradation stems from the low selectivity of certain attributes in the last 3 distributions, which increases unnecessary distance computations, thus reducing overall efficiency.

@@

		%\vspace{0.1cm}
		\begin{subfigure}[t]{0.332\columnwidth}
			\centering
			\captionsetup{font=small}
			\includegraphics[width=\linewidth]{figures/exp/attribute_multimodel.pdf}
			\caption{\footnotesize MultiModel}
			\label{fig:MultiModel}
		\end{subfigure}
		\hfill
		\begin{subfigure}[t]{0.315\columnwidth}
			\centering
			\includegraphics[width=\linewidth]{figures/exp/attribute_85.pdf}
			\caption{\footnotesize Different Computer}
			\label{fig:Different Computer}
		\end{subfigure}
		\hfill
		\begin{subfigure}[t]{0.315\columnwidth}
			\centering
			\includegraphics[width=\linewidth]{figures/exp/attribute_100M.pdf}
			\caption{\footnotesize Big Dataset}
			\label{fig:Big Dataset}
		\end{subfigure}

@@

		%\vspace{0.1cm}
		\begin{subfigure}[t]{0.332\columnwidth}
			\centering
			\captionsetup{font=small}
			\includegraphics[width=\linewidth]{figures/exp/range_multimodel.pdf}
			\caption{\footnotesize MultiModel}
			\label{fig:MultiModel range}
		\end{subfigure}
		\hfill
		\begin{subfigure}[t]{0.315\columnwidth}
			\centering
			\includegraphics[width=\linewidth]{figures/exp/range_85.pdf}
			\caption{\footnotesize Different Computer}
			\label{fig:Different Computer range}
		\end{subfigure}
		\hfill
		\begin{subfigure}[t]{0.315\columnwidth}
			\centering
			\includegraphics[width=\linewidth]{figures/exp/range_deep100M.pdf}
			\caption{\footnotesize Big Dataset}
			\label{fig:Big Dataset range}
		\end{subfigure}
